---
title: "Project 9.2 R Markdown and Leaflet"
author: "ElenaKush"
date: '2020-04-15'
output: html_document
---

## Summary

In this project we will create a web page using R Markdown that features a map created with Leaflet.  
This web page is hosted on GitHub Pages (HERE URL!!!).  

For the project purposes we will analise coronavirus COVID-19 confirmed infection and death cases worldwide.  
We used data posted by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University here: https://github.com/CSSEGISandData/COVID-19.  
The data is incremental and dated from the 2020-01-22 up to the `lastdate`. 

To create a map we used the following libraries: dplyr, leaflet, maps, geojsonio and stringr.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r environment, include = FALSE}
if("pacman" %in% rownames(installed.packages()) == FALSE){ install.packages("pacman") }
library(pacman)
p_load(dplyr,
       leaflet,
       maps,
       geojsonio, 
       stringr)
```

```{r datasets, include = FALSE}
local_path <- "C:/Users/kushnarevaei/Documents/GitHub/JH_R_Course/Course9/Week2/Project9.2/"
URL_cases_global <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"

download.file(URL_cases_global,
              destfile = paste0(local_path,"data/cases_global.csv"))

cases_global <- read.csv(paste0(local_path,"data/cases_global.csv"))

URL_deaths_global <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"

download.file(URL_deaths_global,
              destfile = paste0(local_path,"data/deaths_global.csv"))

deaths_global <- read.csv(paste0(local_path,"data/deaths_global.csv"))

URLfilegeo <- "https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json"

download.file(URLfilegeo,
              destfile = paste0(local_path,"data/countries.geo.json"))

WorldCountry <- geojson_read(paste0(local_path,"data/countries.geo.json"), what = "sp")
```

```{r EDA, include = FALSE}
## World map with countries boundaries:
##View(WorldCountry)
##a <- WorldCountry[!(WorldCountry$name %in% cases_global$Country.Region), ] ##there are old country names whuch do not correspond to the cases_global dataset 
##a[[2]]
correctnames <- c("Antarctica", 
                  "French Southern and Antarctic Lands", 
                  "Bahamas", 
                  "Bermuda",
                  "Cote d'Ivoire",
                  "Congo (Kinshasa)",
                  "Congo (Brazzaville)",
                  "Northern Cyprus",
                  "Czechia",
                  "Falkland Islands",
                  "Guinea-Bissau",
                  "Greenland",
                  "France",
                  "Korea, South",
                  "Lesotho",
                  "North Macedonia",
                  "Burma",
                  "France",
                  "Puerto Rico",
                  "North Korea",
                  "Solomon Islands",
                  "Somalia",
                  "Serbia",
                  "Eswatini",
                  "Tajikistan",
                  "Turkmenistan",
                  "Timor-Leste",
                  "Taiwan*",
                  "Tanzania",
                  "US",
                  "Vanuatu",
                  "West Bank and Gaza")

WorldCountry[!(WorldCountry$name %in% cases_global$Country.Region), 2] <- correctnames  
##dataMap <- WorldCountry[WorldCountry$name %in% cases_global$Country.Region, ]

##head(cases_global)
##View(cases_global)
##str(cases_global)
##dim(cases_global)
##names(cases_global)
lastdate <- names(cases_global)[ncol(cases_global)]
lastdate <- as.Date(str_remove(lastdate, "X"), "%m.%e.%y")

cases_global$total_cases <- cases_global[, ncol(cases_global)]
cases_global <- cases_global %>%
    group_by(Country.Region = cases_global$Country.Region) %>%
    summarise(total_cases = sum(total_cases))



dataMap <- sp::merge(WorldCountry, cases_global, by.x = "name", by.y = "Country.Region", all = T)
##names(dataMap)
dataMap@data[is.na(dataMap@data$total_cases), 3] <- 0
##head(dataMap@data[dataMap@data$total_cases == 0, ])

deaths_global$total_deaths <- deaths_global[, ncol(deaths_global)]
deaths_global <- deaths_global %>%
    group_by(Country.Region = deaths_global$Country.Region) %>%
    summarise(total_deaths = sum(total_deaths))
dataMap <- sp::merge(dataMap, deaths_global, by.x = "name", by.y = "Country.Region", all = T)
##names(dataMap)
dataMap@data[is.na(dataMap@data$total_deaths), 4] <- 0

##
```


```{r viz, include = FALSE}

##head(dataMap)
pal_cases <- colorNumeric(
    palette = 'OrRd',##"Blues",
    domain = as.numeric(dataMap$total_cases))

pal_deaths <- colorNumeric(
    palette = "Blues",
    domain = as.numeric(dataMap$total_deaths))

labels <- sprintf(
    "<strong>Country: %s</strong><br/>Cases: %g<br/>Deaths: %g",
    dataMap$name, dataMap$total_cases, dataMap$total_deaths) %>% 
    lapply(htmltools::HTML) 

leaflet(dataMap) %>% 
    addTiles(group = "COVID19 Confirmed Cases") %>% 
    addTiles(group="COVID19 Deaths") %>%
    addPolygons(
        fillColor = ~ pal_cases(as.numeric(dataMap$total_cases)),
        ##smoothFactor = 0.2,
        weight = 1,
        opacity = 1,
        color = 'gray',
        dashArray = '3',
        fillOpacity = 0.7,
        label = labels,
        highlight = highlightOptions(
            weight = 2,
            color = "#666",
            dashArray = "",
            fillOpacity = 0.7,
            bringToFront = TRUE),
        labelOptions = labelOptions(
            style = list("font-weight" = "normal", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto"),
        group = "COVID19 Confirmed Cases") %>%
    addPolygons(
        fillColor = ~ pal_deaths(as.numeric(dataMap$total_deaths)),
        ##smoothFactor = 0.2,
        weight = 1,
        opacity = 1,
          color = 'gray',
        dashArray = '3',
        fillOpacity = 0.7,
        label = labels,
        highlight = highlightOptions(
            weight = 2,
            color = "#666",
            dashArray = "",
            fillOpacity = 0.7,
            bringToFront = TRUE),
        labelOptions = labelOptions(
            style = list("font-weight" = "normal", padding = "3px 8px"),
            textsize = "15px",
            direction = "auto"),
        group = "COVID19 Deaths") %>%
    addLayersControl(
        baseGroups = c("COVID19 Confirmed Cases", "COVID19 Deaths"),
        options = layersControlOptions(collapsed = FALSE)
  )
    
    
##    addLegend(
  ##    pal = pal, 
    ##  values = ~ as.numeric(dataMap$total_cases), 
      ##opacity = 0.7, 
      ##title = NULL,
      ##position = "topright")

```


